DISCLAIMER: MIS and DIS are identical. In the code the DIS is sometimes referred to as MIS

packages
```{r}
library(tidyverse)
library(sf)
library(spdep)      # Diagnosing spatial dependence
library(spatialreg) # Spatial lag and spatial error model

require(latticeExtra) # USCancerRates, mapplot
require(maps) # map
require(classInt) # classIntervals, findCols
require(grid) # viewport, pushViewport
require(pals) # brewer.blues, stevens.pinkgreen

require("biscale")
require("cowplot")
require("gridExtra")
require(stargazer) # reporting model outputs
library(envalysis)
```

Data
```{r}
# read data set
impacts_year <- readRDS("impacts.rds") %>% select(-c(type_of_class, triple_id, MIS_class, month_date))

# manipulate data to nuts3 and delete 2022
impacts_year <- filter(impacts_year, statistical_unit == "nuts3")
impacts_year <- filter(impacts_year, year_date < 2022)
impacts_year <- distinct(impacts_year, year_nuts, .keep_all = TRUE)

# manipulate income variable
impacts_year <- impacts_year %>%  #group by year
  group_by(year_date)

impacts_year <- impacts_year %>%  #scale income
  mutate(income_scaled = scale(income))

impacts_year$income_scaled <- as.vector(impacts_year$income_scaled) #change from matrix to vector format

impacts_year <- impacts_year %>% #log income
  mutate(income_log = log(income))

impacts_year <- impacts_year %>% #log & scale income
  mutate(income_scaled_log = scale(log(income)))

impacts_year$income_scaled_log <- as.vector(impacts_year$income_scaled_log) # matrix to vector

impacts_year <- ungroup(impacts_year) #ungroup

impacts_year$year_date <- as.factor(impacts_year$year_date) #year_date to factor
```

Heteroscedacity of income, not in the paper
```{r}
log_1 <- ggplot() +
  geom_histogram(filter(impacts_year, year_date %in% c(2003, 2015, 2018, 2019, 2020)), mapping = aes(x = income_log)) +
  facet_wrap(~year_date)

log_2 <- ggplot() +
  geom_histogram(filter(impacts_year, year_date %in% c(2003, 2015, 2018, 2019, 2020)), mapping = aes(x = income_scaled)) +
  facet_wrap(~year_date)

log <- plot_grid(log_1, log_2, 
                 labels = c("log income", "scaled income"))

ggsave("income.jpg", width = 210, height = 150, units = "mm")

log_3 <- ggplot() +
  geom_histogram(filter(impacts_year, year_date %in% c(2003, 2015, 2018, 2019, 2020)),
                 mapping = aes(x = income_scaled_log)) +
  facet_wrap(~year_date)
log_3

log <- plot_grid(log_1, log_2, log_3,
                 labels = c("log income", "scaled income", "log + scaled income"))

ggsave("income_logscaled.jpg", width = 210, height = 150, units = "mm")
```

Datasets Income & education
```{r}
# I create separate datasets for analysis of income and education, I also create the neighbours lists and weights needed for spacial regression analysis

impacts_inc <- filter(impacts_year, year_date %in% c(2003, 2015, 2018, 2019, 2020))

nb_inc <- poly2nb(impacts_inc$geometry, queen = TRUE)
lw_inc <- nb2listw(nb_inc, style = "W", zero.policy = TRUE)


impacts_edu <- filter(impacts_year, year_date %in% c(2015, 2018, 2019, 2020))

nb_edu <- poly2nb(impacts_edu$geometry, queen = TRUE)
lw_edu <- nb2listw(nb_edu, style = "W", zero.policy = TRUE)
```

Spatial autocorrelation of DIS
```{r}
moran.test(impacts_inc$MIS_year, lw_inc)

# There is spatial autocorrelation!
```

Choropleth Plots Income / SPI
```{r}
# SPI

# create legend
legend <- bi_legend(pal = rev("BlueGold"),
                    dim = 3,
                    xlab = "Higher scaled income",
                    ylab = "Higher SPI",
                    size = 12,
                    flip_axes = TRUE)

# loop to create the plots for the individual years
pList <- list()

for (i in c(2003, 2018, 2015, 2019, 2020)) {
  bi_inc <- filter(impacts_inc, year_date %in% c(i))
  bi_inc <- bi_class(bi_inc, y = income_scaled, x = spi_year, style = "quantile", dim = 3)
  bi_inc <- st_as_sf(bi_inc)
map <- ggplot() +
  geom_sf(bi_inc, mapping = aes(fill = bi_class), color = "white", size = 0.1, show.legend = FALSE) +
  bi_scale_fill(pal = "BlueGold", dim = 3) +
  bi_theme()
print(map)
   pList[[i]] <- map
}

# combine the plots in a plot grid and save it
spi_inc_choropleth <- 
  plot_grid(pList[[2003]], pList[[2015]], legend, pList[[2018]], pList[[2019]], pList[[2020]],
            labels = c('2003', "2015", "Legend", '2018', "2019", "2020"))

spi_inc_choropleth

ggsave(file="spi_choropleth_inc.jpg", width = 210, height = 150, units = "mm")

# correlation
cor(impacts_inc$income_scaled, impacts_inc$spi_year)
```

Choropleth Plots Income / DIS
```{r}
# create legend
legend <- bi_legend(pal = "PurpleOr",
                    dim = 3,
                    xlab = "Higher scaled Income",
                    ylab = "Higher DIS",
                    size = 8,
                    flip_axes = TRUE) 

pList <- list()

# as the DIS has a lot of 0 in 2015, the quartiles are not created properly for the plot. Therefore I change all the 0 by a very little bit so that the quartiles can be computed
impacts_inc$MIS_year[impacts_inc$MIS_year == 0 & 
                       impacts_inc$year_date == 2015] <-
  runif(179, 0, 0.00000000000001)

# creating the plots for the years
for (i in c(2003, 2018, 2015, 2019, 2020)) {
  bi_inc <- filter(impacts_inc, year_date %in% c(i))
  bi_inc <- bi_class(bi_inc, y = income_scaled, x = MIS_year, style = "quantile", dim = 3)
  bi_inc <- st_as_sf(bi_inc)
map <- ggplot() +
  geom_sf(bi_inc, mapping = aes(fill = bi_class), color = "white", size = 0.1, show.legend = FALSE) +
  bi_scale_fill(pal = "PurpleOr", dim = 3) +
  bi_theme()
print(map)
   pList[[i]] <- map
}

# combining plots to one
DIS_inc_choropleth <- 
  plot_grid(pList[[2003]], pList[[2015]], legend, pList[[2018]], pList[[2019]], pList[[2020]],
            labels = c('2003', "2015", "Legend", '2018', "2019", "2020"))

DIS_inc_choropleth

ggsave(file="DIS_inc_choropleth.jpg", width = 210, height = 150, units = "mm")

cor(impacts_inc$income_scaled, impacts_inc$MIS_year)
```

Choropleth Plots Education
```{r}
### for SPI
legend <- bi_legend(pal = rev("BlueGold"),
                    dim = 3,
                    xlab = "Higher education",
                    ylab = "Higher SPI",
                    size = 12,
                    flip_axes = TRUE) 

legend

for (i in c(2015, 2018, 2019, 2020)) {
  bi_edu <- filter(impacts_edu, year_date %in% c(i))
  bi_edu <- bi_class(bi_edu, y = education, x = spi_year, style = "quantile", dim = 3)
  bi_edu <- st_as_sf(bi_edu)
map <- ggplot() +
  geom_sf(bi_edu, mapping = aes(fill = bi_class), color = "white", size = 0.1, show.legend = FALSE) +
  bi_scale_fill(pal = "BlueGold", dim = 3) +
  bi_theme()
print(map)
   pList[[i]] <- map
}

SPI_choropleth <- 
  plot_grid( "", pList[[2015]],legend, pList[[2018]], pList[[2019]], pList[[2020]],
            labels = c("","2015", "Legend", '2018', "2019", "2020"))

SPI_choropleth

ggsave(file="SPI_choropleth_edu.jpg", width = 210, height = 150, units = "mm")

cor(impacts_edu$spi_year, impacts_edu$education)
```


```{r}
### for DIS
legend <- bi_legend(pal = "PurpleOr",
                    dim = 3,
                    xlab = "Higher education",
                    ylab = "Higher DIS",
                    size = 8,
                    flip_axes = TRUE) 

legend

pList <- list()

impacts_edu$MIS_year[impacts_edu$MIS_year == 0 & 
                       impacts_edu$year_date == 2015] <-
  runif(179, 0, 0.00000000000001)

for (i in c(2015, 2018, 2019, 2020)) {
  bi_edu <- filter(impacts_edu, year_date %in% c(i))
  bi_edu <- bi_class(bi_edu, y = education, x = MIS_year, style = "quantile", dim = 3)
  bi_edu <- st_as_sf(bi_edu)
map <- ggplot() +
  geom_sf(bi_edu, mapping = aes(fill = bi_class), color = "white", size = 0.1, show.legend = FALSE) +
  bi_scale_fill(pal = "PurpleOr", dim = 3) +
  bi_theme()
print(map)
   pList[[i]] <- map
}

DIS_choropleth <- 
  plot_grid("", pList[[2015]],legend, pList[[2018]], pList[[2019]], pList[[2020]],
            labels = c("", "2015", "Legend", '2018', "2019", "2020"))

DIS_choropleth

ggsave(file="MIS_choropleth_edu.jpg", width = 210, height = 150, units = "mm")

cor(impacts_edu$MIS_year, impacts_edu$education)
```

Spatial autocorrelation
```{r}
DIS.lm <- 
  lm(MIS_year ~
       log(income) 
     + spi_year + east + agriculture + forest + settlement_transportation + water_bodies + as.factor(year_date) + as.factor(year_date)*log(income) + east*log(income), 
data = impacts_inc
     )

DIS.lm %>% summary()

impacts_inc$residuals <- residuals(DIS.lm)
impacts_inc$fitted <- fitted(DIS.lm)

lm.morantest(DIS.lm, lw, zero.policy = T) #spatial effect is there, Moran I is significant

lm.LMtests(DIS.lm, lw, test="LMerr", zero.policy = T) #spatial error model: significant

lm.LMtests(DIS.lm, lw, test="LMlag", zero.policy = T) #spatial lag model: significant

moran.plot(impacts_inc$residuals, lw, zero.policy = TRUE) #plot spatiall lagged values

moran <- moran.mc(impacts_inc$residuals, lw, nsim = 999, zero.policy = TRUE)
moran

## sensetivity
 impacts_inc %>% 
  mutate(lon = list(st_coordinates(.)[, 1]),
            lat = list(st_coordinates(.)[, 2])) %>% 
  unnest(lon, lat) %>% 
  st_drop_geometry()
 
 moran_I <- c()
 
 for (d in seq(50, 2000, 50)) {
  nb <- dnearneigh(impacts_inc$geometry, d1 = 0, d2 = d)
  lw <- nb2listw(nb, style = "W", zero.policy = TRUE)
  moran <- moran.mc(DIS.lm$residuals, lw, nsim = 999, zero.policy = TRUE)
  moran_I <- c(moran_I, moran$statistic)
 }
 
 #spatial lag model
```

Hypothethis 2: income
```{r}
# 4 spatial error models! 

# just DIS and income
model_1 <- errorsarlm(MIS_year ~ income_scaled, impacts_inc, lw_inc)

# DIS, income and control
model_2 <- errorsarlm(MIS_year ~ income_scaled + spi_year + agriculture + forest + settlement_transportation + water_bodies + year_date + east, impacts_inc, lw_inc)

# DIS, income, control and interaction term with east for H2b
model_3 <- errorsarlm(
  MIS_year ~ income_scaled + spi_year + agriculture + forest + settlement_transportation + water_bodies + year_date + east + east*income_scaled,
  impacts_inc,
  listw = lw_inc,
  zero.policy = TRUE)

# DIS, income, control and interaction with agiculture for H2c
model_4 <- errorsarlm(
  MIS_year ~ income_scaled + spi_year + agriculture + forest + settlement_transportation + water_bodies + year_date + east + income_scaled*agriculture,
  impacts_inc,
  listw = lw_inc,
  zero.policy = TRUE)
```

Hypothethis 2: eduaction
```{r}
# model DIS, education
model_1_edu <- errorsarlm(MIS_year ~ education, impacts_edu, lw_edu)

# model DIS, education and control
model_2_edu <- errorsarlm(MIS_year ~ education + spi_year + agriculture + forest + settlement_transportation + water_bodies + year_date + east, impacts_edu, lw_edu)

# labels for stargazer
labels_1 <- c("Income scaled", 
                               "Education",
                               "SPI",
                               "Agriculture",
                               "Forestry",
                               "Settlement + Transport",
                               "Water",
                               "2015",
              "2018",
              "2019",
              "2020",
              "East",
              "Income:East",
              "Income:Agriculture")

# stargazer output
stargazer(model_1, model_1_edu, model_2, model_2_edu, model_3, model_4,
          out = "model_all.htm",
          column.sep.width = "10pt",
          covariate.labels = labels_1,
          dep.var.labels = "DIS")
```

Homoscedacity of residuals, not in the work included
```{r}
test <- data.frame(
  fitted = model_2_edu$fitted.values,
  resid = model_2_edu$residuals
)

ggplot(test) +
  geom_point(aes(x = fitted, y = resid))

bptest.Sarlm(model_2_edu)

# heteroscedacity is there...
```

Hypothesis 3
```{r}
# change dataset to only include 2018 - 2020 and create neighbours and weights 
impacts_3 <- impacts_year %>%
  ungroup() %>% 
  filter(year_date %in% c("2018" ,"2019" ,"2020")) %>% 
  select(nuts_id, MIS_year, spi_year, education, income_scaled, year_date, 
         geometry, east) %>% 
  gather("key", "value", 2:5) %>% 
  unite("year_key", key, year_date) %>% 
  spread(year_key, value)

nb_3 <- poly2nb(impacts_3$geometry, queen = TRUE)
lw_3 <- nb2listw(nb_3, style = "W", zero.policy = TRUE)

#calculate how DIS has changed over the time

impacts_3$DIS_diff <- impacts_3$MIS_year_2020 - impacts_3$MIS_year_2018
impacts_3$DIS_diff19 <- impacts_3$MIS_year_2019 - impacts_3$MIS_year_2018

# calculate how SPI has changed over time

impacts_3$SPI_diff <- impacts_3$spi_year_2020 - impacts_3$spi_year_2018
impacts_3$SPI_diff19 <- impacts_3$spi_year_2019 - impacts_3$spi_year_2018

# create plots of how DIS has changed relaitve to income / education

diff_1 <- ggplot(impacts_3) +
  geom_point(aes(x = DIS_diff19, y = income_scaled_2018)) +
  ylab("Income scaled 2018") +
  xlab(NULL)+
  theme_publish()+
  theme(plot.title.position = "plot")

diff_2 <- ggplot(impacts_3) +
  geom_point(aes(x = DIS_diff, y = income_scaled_2018)) +
  ylab(NULL) +
  xlab(NULL) +
  theme_publish()+
  theme(plot.title.position = "plot")

diff_3 <- ggplot(impacts_3) +
  geom_point(aes(x = DIS_diff19, y = education_2018)) +
  ylab("Education 2018") +
  xlab("Development DIS 2018 - 2019") +
  theme_publish()+
  theme(plot.title.position = "plot")

diff_4 <- ggplot(impacts_3) +
  geom_point(aes(x = DIS_diff, y = education_2018)) +
  xlab("Development DIS 2018 - 2020") +
  ylab(NULL) +
  theme_publish()+
  theme(plot.title.position = "plot")

diff <- plot_grid(diff_1, diff_2, diff_3, diff_4)
diff

ggsave("Differences_plot.jpg", width = 210, height = 150, units = "mm")
```

Model Hyp 3
```{r}
# four models for H3

# income and change to 2020
model_diff_inc <- errorsarlm(DIS_diff ~ income_scaled_2018 + SPI_diff + east, impacts_3, lw_3)

# income and change to 2019
model_diff_inc_19 <- errorsarlm(DIS_diff19 ~ income_scaled_2018 + SPI_diff + east, impacts_3, lw_3)

# education and change to 2020
model_diff_ed <- errorsarlm(DIS_diff ~ education_2018 + SPI_diff + east, impacts_3, lw_3)

# education and change to 2019
model_diff_ed_19 <- errorsarlm(DIS_diff19 ~ education_2018 + SPI_diff + east, impacts_3, lw_3)

# labels for stargazer and stargazer output
labels_2 <- c("Income scaled 2018",
              "Education 2018",
              "Development SPI",
              "East")

stargazer(model_diff_inc_19, model_diff_inc, model_diff_ed_19 ,model_diff_ed, 
          out = "diff.htm", 
          single.row = TRUE,
          covariate.labels = labels_2,
          dep.var.labels = c("DIS to 2019",
                             "DIS to 2020",
                             "DIS to 2019",
                             "DIS to 2020")
          )
```

